G2GML: Graph to Graph Mapping Language for Bridging RDF and Property Graphs

### 2020-04-18 paper.tex
Update after English editing

\begin{document}
\newtheorem{defi}[theorem]{Definition}
\newtheorem*{prel}{Preliminaries}
%
\title{G2GML: Graph to Graph Mapping Language for Bridging RDF and Property Graphs}
%
\titlerunning{G2GML: Graph to Graph Mapping Lanugage}
%\titlerunning{Interoperable Use of RDF and Property Graph}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Hirokazu Chiba\inst{1} \and Ryota Yamanaka\inst{2} \and Shota Matsumoto\inst{3}}
%
\authorrunning{H. Chiba {\itshape et al.}}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
Database Center for Life Science, Chiba 277-0871, Japan\\
\email{chiba@dbcls.rois.ac.jp}
\and
Oracle Corporation, Bangkok 10500, Thailand\\
\email{ryota.yamanaka@oracle.com}
\and
Lifematics Inc., Tokyo 101-0041, Japan\\
\email{shota.matsumoto@lifematics.co.jp}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
% The abstract should briefly summarize the contents of the paper in 15--250 words.
How can we maximize the value of accumulated RDF data? Although increasing amounts of scientific and social data are described as RDF, their use cases are still limited. Allowing access to the RDF datasets through various interface beyond the SPARQL interface will be expected. Importing the RDF datasets into various graph analysis engines will increase the values of the datasets.  
However, interoperable management of such graph data is challenging due to the differences in existing models and formats depending on various implementations. Here, we redefine the property graph model incorporating the differences in existing models and propose exchangeable serialization formats for property graphs, which can be converted into specific formats for each of the databases. 
The serialization formats independent of certain database implementations will increase the interoperability of graph databases and will make it easier for users to import accumulated graph data.
We further propose a framework for mapping RDF graphs to property graphs on the basis of the Graph to Graph Mapping Language (G2GML). Using this framework, accumulated graph data described in RDF can be converted to property graphs, which can then be loaded into several graph database engines for further analysis. 
%This work provides interoperability not only between property graph formats but also between RDF and proper graphs.
%Future works include implementing and utilizing graph algorithms to make the most of the accumulated data in various analytical engines.

@@

\section{Introduction}
% Please note that the first paragraph of a section or subsection is not indented. The first paragraph that follows a table, figure, equation etc. does not need an indent, either.

Increasing amounts of scientific and social data are published in the form of Resource Description Framework (RDF)~\cite{rdf}. It now constitutes a large open data cloud. DBpedia~\cite{dbpedia} and Wikidata~\cite{wikidata} are good example. SPARQL~\cite{sparql} is a standardized interface for RDF. However, the use cases through thouse interface are still limited. Although the standardized format and interface contributed the integrated large data could, further interface is expected to maximize the value of RDF datasets. 
% Although RDF data can be queried using the SPARQL language~\cite{sparql} in a flexible way, SPARQL is not dedicated to traversal of graphs and has a limitation in implementing graph analysis algorithms.
Recently, the property graph model~\cite{angles1,angles2} have been attracting increasing attention in the context of graph analysis; various graph database engines, including Neo4j~\cite{neo4j}, Oracle Database~\cite{oracle}, and Amazon Neptune~\cite{neptune}, adopt this model. These graph database engines support algorithms for traversing or analyzing graphs. However, currently not many datasets are consistently described in the property graph model, so the application of these powerful engines are limited.

Considering this situation, it is valuable to develop a method to transform RDF into property graphs. One of the practical issues is the lack of a standardized model for property graphs. 
Another issue is that transformation between them is not straightforward due to the differences between RDF and property graph models. 
In RDF graphs, all information is expressed as the triple (node-edge-node), whereas in property graphs, arbitrary information can be contained in each of the nodes and edges as the key-value form. 
Although this issue has been previously addressed on the basis of predefined transformations~\cite{hartig},
users still cannot control the mapping for their specific use cases.

Here, we redefine the property graph model incorporating the differences in existing models and also propose serialization formats based on the data model. We further propose a graph to graph mapping framework on the basis of the Graph to Graph Mapping Language (G2GML). Using this mapping framework, accumulated graph data described in RDF can be converted into property graphs, which can then be loaded into several graph database engines for further analysis. We will show several use cases where publicly available RDF data is extracted and convert to property graphs and also evaluate the performance. This work provides the foundation of interoperability of knowledge graphs.

@@

The main contributions of this papers are 1) language design of G2GML, 2) its prototype implementation. Additionally, we proposes 3) the common property graph model and its serialization, which are essential to make this mapping framework independent from the implementations of database systems.

@@

Here we illustrate the overview of the graph to graph mapping (G2G mapping) framework (Figure~\ref{fig:dataflow}).

In this framework, users describe mappings from RDF to property graphs in G2GML.
According to this G2GML description, a G2GML implementation converts input RDF dataset into property graph dataset. Optionally, the new dataset can be also saved in specific formats for loading it into the major graph database implementations.

%G2G is a framework to perform mapping between RDF and property graphs for interoperable use of them. In this framework, users can describe mappings from RDF to property graphs in G2GML.
%G2GML implementation converts RDF dataset into property graph dataset. Optionally, the new dataset can be also saved in specific formats for loading it into the major graph database implementations.

The G2GML is a declarative language which consists of pairs of RDF graph patterns and property graph patterns. 
An intuitive meaning of a G2GML description is a map from RDF subgraphs that matches the specified patterns to specified components of property graphs.

As an entry point to property graph, we define the Property Graph format, which is based on the Property Graph model.

\begin{figure}
\center
\includegraphics[width=1.0\textwidth]{dataflow.png}
\caption{Overview of the mapping from RDF to property graph}
\label{fig:dataflow}
\end{figure}

@@

G2GML is a language to describes maps from RDF graphs to property graphs. G2GML is a domain specific declarative language where programmers only have to describe patterns of RDF and property graphs (PG). RDF pattern is described using SPARQL pattern, and PG pattern is described using Cypher pattern.

\begin{prel}
An RDF triple $(s, p, o)$ is an element of $(I \cup B) \times I \times (I \cup B \cup L)$ where $I$, $L$ and $B$ are a set of IRIs, a set of literals and a set of blank nodes, which are considered pairwise disjoint. In this paper, an RDF triple is simply called a triple. For a triple $(s, p, o)$, $s$ is called the subject, $p$ the predicate and $o$ the object. An RDF graph is defined as a finite set of triples.
\end{prel}

\par
A map described by G2GML consists of one or more \textit{node maps} and \textit{edge maps}.
A \textit{node map} is a function that maps from an RDF graph to a set of \textit{PG nodes}. 
Similarly, an \textit{edge map} is a function that maps from an RDF graph to a set of \textit{PG edges}.
The meaning of a G2GML description is a function that maps from an RDF graph to a set of PG nodes and edges that is mapped by all its node and edge maps.
Here we define the syntax of PG node pattern as follows. 
Note that \texttt{NEWLINE} is an abstract symbol corresponding to appropriate character to start a new line.

\begin{defi}[EBNF notation of G2GML]
\label{defi:G2GML}
\leavevmode
\begin{verbatim}
 g2gml   ::= SPARQL_PREFIX* mapping+
 mapping ::= ( node_pat | edge_pat ) NEWLINE
               INDENTED_RDF_PATTERN NEWLINE
\end{verbatim}
\end{defi}
In Definition \ref{defi:G2GML}, \texttt{SPARQL\_PREFIX}
 and \texttt{INDENTED\_RDF\_PATTERN} are almost same as \textit{PrefixDecl} and \textit{GroupGraphPattern} in SPARQL syntax (see EBNF notation of SPARQL grammer~\cite{sparql}).
The only difference is that \texttt{INDENTED\_RDF\_PATTERN} must have one or more white spaces at the top of each line.
Symbols \texttt{node\_pat} \texttt{edge\_pat} will be described in the following definition.

An RDF graph pattern in G2GML specifies which resources should be mapped to PG nodes. Resources matching with these pattern yield solution sets of SELECT query in SPARQL, which 
 are mapped to PG nodes by embedding variable bindings to corresponding location according to the preceding PG node pattern.
A complete example of G2GML is shown in Figure \ref{fig:example-g2g}.
%TODO: consider the position of the example
A node map is described as a pair of a PG node pattern  and an RDF graph pattern. Here we define the syntax of PG node pattern as follows. Note that \texttt{NODE}, \texttt{LABEL}, \texttt{PROP\_NAME} and \texttt{PROP\_VAL} are all identifiers.
\begin{defi}[EBNF notation of PG node pattern]
\leavevmode
\begin{verbatim}
 node_pat  ::= "(" NODE  ":" LABEL  ("{" prop_list "}")? ")"
 prop_list ::= property "," prop_list | property
 property  ::= PROP_NAME ":" PROP_VAL
\end{verbatim}
\end{defi}

A PG node pattern specifies a node, a label and properties to be generated by the map. The \texttt{NODE} symbol specifies ID of the generated node, which is replaced by the resource retrieved by RDF graph. The \texttt{LABEL} symbol specifies the label of the resultant node. A label also works as the name of the node map which is referred from edge maps.
Therefore an identifier of a label must be unique in a G2GML description.
Each node pattern contains zero or more properties as \texttt{prop\_list}.
A \texttt{property} is a pair of \texttt{PROP\_NAME} and \texttt{PROP\_VAL} that describes properties of nodes.
Identifiers \texttt{NODE} and \texttt{PROP\_VAL} must be names of variables in the succeeding RDF graph pattern, while \texttt{LABEL} and \texttt{PROP\_NAME} are regarded as text literals.

A PG edge pattern specifies a pattern of edge, the label, the properties and the source node and the destination node.
Here we define the syntax of PG edge pattern as follows.
Note that \texttt{SRC}, \texttt{SRC\_LAB}, \texttt{EDGE}, \texttt{EDGE\_LAB}, \texttt{DST} and \texttt{DST\_LAB} are all identifiers.
\begin{defi}[EBNF notation of PG edge pattern]
\leavevmode
\begin{verbatim}
edge_pat ::= "(" SRC  ":" SRC_LAB ") -" 
                "[" EDGE?  ":" EDGE_LAB  ("{" prop_list "}")? ")" 
                ("->" | "-") "(" DST ":" DST_LAB ")"
\end{verbatim}
\end{defi}

 Identifiers of \texttt{SRC} and \texttt{DST} specify which variables in the succeeding RDF graph pattern should be mapped to endpoint nodes of resultant edges.
A \texttt{SRC\_LAB} and a \texttt{DST\_LAB} work not only as labels of resultant nodes but also work as implicit constraints of the edge map.
%TODO add ref to example section.
The resultant source nodes and destination nodes must match patterns in corresponding node maps.
For this reason, \texttt{SRC\_LAB} and \texttt{DST\_LAB} must be defined in other node maps.
\texttt{EDGE}, \texttt{EDGE\_LAB} and \texttt{prop\_list} can be described in the same manner as those counterparts in node maps.

The resultant edges can be either directed or undirected. If we use '\texttt{->}' as delimiter between an edge part and a destination part, the edge will be a directed edge.

In node maps and edge maps, a variable bound to \texttt{PROP\_VAL} may have multiple candidates against single node or edge. If multiple candidate exists, they are concatenated into an array in the manner of GROUP\_CONCAT in SPARQL. If \texttt{EDGE} is omitted in an edge map, such candidates are concatenated  within groups of tuples of \texttt{SRC} and \texttt{DST}.

%%%%%%%%%%%
\subsection{The Property Graph Model}
Here we define the property graph model independent of specific graph database implementations. For the purpose of interoperability, we incorporate differences in property graph models, taking into consideration multiple labels or property values for nodes and edges, as well as mixed graphs with both of directed and undirected edges. The property graph model we redefine here requires the following characteristics:

\begin{itemize}
    \item A property graph contains nodes and edges.
    \item Each of the nodes and edges can have zero or more labels.
    \item Each of the nodes and edges can have properties (key-value pairs).
    \item Each property can have multiple values.
    \item Each edge can be directed or undirected.
\end{itemize}
More formally, we define the property graph model as follows.

\begin{defi}[Property Graph Model]
\leavevmode \vspace{1mm} \\
A \emph{Property Graph} is a tuple
$PG = \langle N, E_d, E_u, S, V, P, e, l_n, l_e, p_n, p_e\rangle$, where:
\begin{enumerate}
    \item $N$ is a set of nodes.
    \item $E_d$ is a set of directed edges.
    \item $E_u$ is a set of undirected edges.
    \item $E$ is a set of edges where $E = E_d \cup E_u$.
    \item $S$ is a set of strings.
    \item $V$ is a set of values of arbitrary data types.
    \item $P$ is a set of properties. Each property has the form $p = \langle k,v \rangle$, where $k \in S$ and $v \in 2^V$.
    \item $e: E \to \langle N \times N \rangle$ is a function which yields the endpoints of each directed or undirected edge (if the edge is directed, the first node is the source and the second node is the destination).
    \item $l_n : N \to 2^S$ is a function mapping each node to its multiple labels.
    \item $l_e : E \to 2^S$ is a function mapping each edge to its multiple labels.
    \item $p_n : N \to 2^P$ is a function used to assign nodes to their multiple properties.
    \item $p_e : E \to 2^P$ is a function used to assign edges to their multiple properties.
\end{enumerate}
\end{defi}

\subsection{Serialization of Property Graphs}
According to our definition of the property graph model, we propose serialization in flat text and JSON. The flat text format (PG format) is better for human readability and line-oriented processing, while the JSON format (JSON-PG format) is best used for server-client communication.

@@

\begin{itemize}
    \item Each line describes a node or an edge.
    \item All elements in each line are separated by space or tab.
    \item In each of the node lines, the first column contains the node ID.
    \item In each of the edge lines, the first three columns contain the source node ID, the direction, and the destination node ID.
    \item Each line can contain an arbitrary number of labels.
    \item Each line can contain an arbitrary number of properties (key-value pairs).
\end{itemize}

@@

% Next, we describe the JSON-PG format which follows the JSON syntax in addition to the above definition of the property graph model. The JSON-PG format has the following characteristics, and an example of the format is shown in Figure~\ref{fig:example-json}. It is to be noted that, whereas the set of labels or property values are represented as arrays in JSON, those elements are supposed to have no specific order according to the property graph model.

% \begin{itemize}
%     \item Nodes and edges are listed under \texttt{nodes} and \texttt{edges} elements, respectively.
%     \item Edge direction is defined with the boolean element \texttt{undirected}. By default it is false (directed).
%     \item Labels are listed under the \texttt{labels} element.
%     \item Properties (key-value pairs) are listed under the \texttt{properties} element.
% \end{itemize}

Furthermore, we have implemented command-line tools to convert between PG format and JSON-PG format which follows the JSON syntax in adition to the above definition, as well as to transform them into formats for well-known graph databases such as Neo4j, Oracle Database, and Amazon Neptune. The practical use cases of our tools demonstrate that the proposed data model and formats have the capability to describe property graph data being used in existing graph databases (see \url{https://github.com/g2glab/pg}).

@@

% \begin{figure}[!t]
% \begin{scriptsize}
% \begin{verbatim}
% {
%   "nodes":[
%     {
%      "id":101,
%      "labels":["Person"],
%      "properties":{"name":["Alice"], "age":[15], "country":["United States"]}
%     },
%     {
%      "id":102,
%      "labels":["Person", "Student"],
%      "properties":{"name":["Bob"], "country":["Japan", "Germany"]}
%     }
%   ],https://ja.overleaf.com/project/5d6e87b11a254100014e487d
%   "edges":[
%     {
%      "from":101,
%      "to":102,
%      "undirected":true,
%      "labels":["sameSchool", "sameClass"],
%      "properties":{"since":[2012]}
%     },
%     {
%      "from":102,
%      "to":101,
%      "labels":["likes"],
%      "properties":{"since":[2015]}
%     }
%   ]
% }
% \end{verbatim}
% \end{scriptsize}
% \caption{Example of JSON-PG format}
% \label{fig:example-json}
% \end{figure}

%%%%%%%%%%%%
\subsection{Minimal Examples}
Figure~\ref{fig:example-g2g} shows the small example of G2G mapping from RDF data to PG data. This example shows the following five types of the typical mapping.

\begin{itemize}
    \item Resource to node: In the line 2-4, the RDF resources with type \texttt{:Person} are mapped into the PG nodes using their IRIs as node IDs.
    \item Datatype property to node property: In the line 2-4, the RDF datatype property \texttt{:name} is mapped into the PG node property key \texttt{name}. The literal objects 'Alice' and 'Bob' are mapped into the node property values.
    \item Object property to edge: In the line 5-6, the RDF object property \texttt{:supervised\_by} is mapped into the PG edge \texttt{supervised\_by}.
    \item Resource to edge: In the line 7-10, the RDF resource with type \texttt{:Email} is mapped into the PG edge \texttt{emailed}. 
    \item Datatype property to edge property: In the line 7-10, the RDF datatype property \texttt{:year} and \texttt{:attachment} are mapped into the PG edge property \texttt{year} and \texttt{attachment}. The literal objects \texttt{2017} and \texttt{01.pdf} are mapped into the edge property values.
\end{itemize}

@@

%%%%%%%%%%%
\subsection{Mapping Details}
The G2G mapping based on G2GML is generally intuitive as it is shown above, however, there are several discussions about the details of mapping.

\subsubsection{Referencing Node Labels}
The node labels in edge maps must be defined in node maps  (described in section 2.2) for referencing the conditions. This means that the conditions for defining `Person` are inherited to the definition of `supervised\_by` and `emailed` relationships. Therefore, when any `supervised\_by` relationship exists between the nodes which are not `Person`, this relationship will not be generated. In this way, users can retrieve the property graph datasets which are intuitively consistent (it is also possible not to specify node labels). On the other hand, edge conditions are not inherited to nodes, so the nodes which do not have any `supervised\_by` nor `emailed` relationships are also generated.

\subsubsection{Multi Edges}
When a pair of RDF resources has multiple relationships, those will be converted not to a single PG edge with multiple properties but to multiple PG edges. The G2GML in Figure~\ref{fig:example-g2g} generates two PG edges for the RDF data in Figure~\ref{fig:example-rdf2} for keeping the information that Alice emailed Bob twice.

\begin{figure}[!t]
\begin{scriptsize}
\begin{verbatim}
@prefix : <http://example.org/> .
:person1 a :Person ;
         :name 'Alice' .
:person2 a :Person ;
         :name 'Bob' .
[] a :Email ;
   :sender   :person1 ;
   :receiver :person2 ;
   :year     2017 .
[] a :Email ;
   :sender   :person1 ;
   :receiver :person2 ;
   :year     2018 .
\end{verbatim}
\end{scriptsize}
\caption{Example of input RDF data (multi edges)}
\label{fig:example-rdf2}
\end{figure}

\begin{figure}[!t]
\begin{scriptsize}
\begin{verbatim}
"http://example.org/person1" :Person name:Alice
"http://example.org/person2" :Person name:Bob
"http://example.org/person1" -> "http://example.org/person2" :emailed year:2017
"http://example.org/person1" -> "http://example.org/person2" :emailed year:2018
\end{verbatim}
\end{scriptsize}
\caption{Example of output PG data (multi edges)}
\label{fig:example-pg2}
\end{figure}

>>>>

\begin{document}
\newtheorem{defi}[theorem]{Definition}
\newtheorem*{prel}{Preliminaries}
%
\title{G2GML: Graph to Graph Mapping Language for Bridging RDF and Property Graphs} 
%
\titlerunning{G2GML: Graph to Graph Mapping Language}
%\titlerunning{Interoperable Use of RDF and Property Graph}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Hirokazu Chiba\inst{1} \and Ryota Yamanaka\inst{2} \and Shota Matsumoto\inst{3}}
%
\authorrunning{H. Chiba {\itshape et al.}}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
Database Center for Life Science, Chiba 277-0871, Japan\\
\email{chiba@dbcls.rois.ac.jp}
\and
Oracle Corporation, Bangkok 10500, Thailand\\
\email{ryota.yamanaka@oracle.com}
\and
Lifematics Inc., Tokyo 101-0041, Japan\\
\email{shota.matsumoto@lifematics.co.jp}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Increasing amounts of scientific and social data are presented in the resource description framework (RDF); however, their use cases remain limited. Maximizing the value of accumulated RDF data is therefore considered a challenge, and efforts have been made to this end, including allowing access to RDF datasets through various interfaces beyond SPARQL and importing RDF datasets into various graph analysis engines.  
However, the interoperable management of graph data is challenging due to differences in existing models and formats, which depend on various implementations. In this study, we redefine the property graph (PG) model incorporating the differences in existing models and propose exchangeable serialization formats for PGs, which can be converted into specific formats for each database. 
The serialization formats that are independent of certain database implementations increase the interoperability of graph databases and make it easier for users to import accumulated graph data.
We further propose a framework for mapping RDF graphs to PGs based on the graph to graph mapping language (G2GML). Using this framework, accumulated graph data described in RDF can be converted into PGs, which can then be loaded into several graph database engines for further analysis.

@@

\section{Introduction}

Increasing amounts of scientific and social data are being published in the form of the Resource Description Framework (RDF)~\cite{rdf}, which presently constitutes a large open data cloud. DBpedia~\cite{dbpedia} and Wikidata~\cite{wikidata} are good examples of such RDF datasets. SPARQL~\cite{sparql} is a protocol and query language that serves as a standardized interface for RDF data. This standardized format and interface enables the construction of integrated graph datasets. However, the lack of an interface for graph-based analysis and performant traversal limits use cases of the graph data.

Recently, the property graph (PG) model~\cite{angles1,angles2} has been attracting increasing attention in the context of graph analysis. Various graph database engines, including Neo4j~\cite{neo4j}, Oracle Database~\cite{oracle}, and Amazon Neptune~\cite{neptune} adopt this model. These graph database engines support algorithms for traversing or analyzing graphs. However, few datasets are consistently described in the PG model, such that the application of these powerful engines is limited.

In light of this situation, developing a method to transform RDF into PG would be highly valuable. One of the practical issues faced by this challenge is the lack of a standardized PG model.
Another issue is that the transformation between RDF and PG is not straightforward due to the differences in their models. 
In RDF graphs, all information is expressed by a triple (node-edge-node), whereas in PGs, arbitrary information can be contained in each node and edge as the key-value form. 
Although this issue was previously addressed on the basis of predefined transformations~\cite{hartig},
users still cannot control the mapping for their specific use cases.

In this study, we redefine the PG model incorporating the differences in existing models and propose serialization formats based on the data model. We further propose a graph to graph mapping framework based on the graph to graph mapping language (G2GML). Employing this mapping framework, accumulated graph data described in the RDF can be converted into PGs, which can be loaded into several graph database engines for further analysis. We demonstrate several use cases, where publicly available RDF data is extracted and converted to PGs, and evaluate the performance. Thus, this study provides the foundation for the interoperability of knowledge graphs.

@@

The main contributions of this study are as follows: 1) language design of G2GML; 2) its prototype implementation. Furthermore, we propose 3) the common PG model and its serialization, which are essential to ensure this mapping framework is independent from the implementations of database systems.

@@

We provide an overview of the graph to graph mapping (G2G mapping) framework (Figure~\ref{fig:dataflow}).

In this framework, users describe mappings from the RDF to PG in G2GML.
According to this G2GML description, G2GML implementation converts the input RDF dataset into the PG dataset. The new dataset can also be optionally saved in specific formats for loading into major graph database implementations.

The G2GML is a declarative language that consists of pairs of RDF patterns and PG patterns. 
The intuitive meaning of a G2GML can be represented by a map of RDF subgraphs, which matches the specified patterns to specified components of PGs.

%As an entry point to the PG, we define serialization based on the PG model.
%G2GML implementation converts the RDF dataset into a property graph dataset. The new dataset can also be saved in specific formats for loading into major graph database implementations.

\begin{figure}
\center
\includegraphics[width=1.0\textwidth]{dataflow.png}
\caption{Overview of mapping from RDF to PG}
\label{fig:dataflow}
\end{figure}

@@

G2GML is a language applied to describe maps from RDF graphs to PGs. It represents a domain-specific declarative language, where programmers only have to describe patterns of RDF and PG. The RDF pattern is described in terms of the SPARQL pattern, and the PG pattern is described using the Cypher pattern.

\begin{prel}
An RDF triple $(s, p, o)$ is an element of $(I \cup B) \times I \times (I \cup B \cup L)$, where $I$, $L$, and $B$ are a set of IRIs, a set of literals, and a set of blank nodes, which are considered pairwise disjoint. In this study, an RDF triple is simply referred to as a triple. For a triple $(s, p, o)$, $s$ is the subject, $p$ the predicate, and $o$ the object. An RDF graph is defined as a finite set of triples.
\end{prel}

\par
A G2GML description consists of one or more \textit{node maps} and \textit{edge maps}.
A \textit{node map} is a function that maps resources from an RDF graph to a set of \textit{PG nodes}. 
Similarly, an \textit{edge map} is a function that maps resources from an RDF graph to a set of \textit{PG edges}.
%A G2GML description is a function that maps resources from an RDF graph to a set of PG nodes and edges that are mapped by all node and edge maps.
Here, we define the syntax of the PG node pattern as follows. 
\texttt{NEWLINE} is used an abstract symbol corresponding to the appropriate character to start a new line.

\begin{defi}[EBNF notation of G2GML]
\label{defi:G2GML}
\leavevmode
\begin{verbatim}
 g2gml   ::= SPARQL_PREFIX* mapping+
 mapping ::= ( node_pat | edge_pat ) NEWLINE
               INDENTED_RDF_PATTERN NEWLINE
\end{verbatim}
\end{defi}
In Definition \ref{defi:G2GML}, \texttt{SPARQL\_PREFIX}
 and \texttt{INDENTED\_RDF\_PATTERN} are almost the same as \textit{PrefixDecl} and \textit{GroupGraphPattern} in SPARQL syntax (see EBNF notation of SPARQL grammar~\cite{sparql}).
The only difference is that \texttt{INDENTED\_RDF\_PATTERN} must have one or more white spaces on top of each line.
Symbols \texttt{node\_pat} \texttt{edge\_pat} are described according to the following definition.

An RDF graph pattern in G2GML specifies the resources that should be mapped to PG nodes. Resources matching with these patterns yield solution sets of SELECT query in SPARQL, which are mapped to PG nodes by embedding variable bindings to corresponding locations according to the preceding PG node pattern.
A complete example of G2GML is shown in Figure \ref{fig:example-g2g}.
%TODO: consider the position of the example
A node map is described as a pair of a PG node pattern and an RDF graph pattern. Here, we define the syntax of the PG node pattern as follows. Notably, \texttt{NODE}, \texttt{LABEL}, \texttt{PROP\_NAME}, and \texttt{PROP\_VAL} are all identifiers.
\begin{defi}[EBNF notation of PG node pattern]
\leavevmode
\begin{verbatim}
 node_pat  ::= "(" NODE  ":" LABEL  ("{" prop_list "}")? ")"
 prop_list ::= property "," prop_list | property
 property  ::= PROP_NAME ":" PROP_VAL
\end{verbatim}
\end{defi}

A PG node pattern specifies a node, a label, and properties to be generated by the map. The \texttt{NODE} symbol specifies the ID of the generated node, which is replaced by the resource retrieved by the RDF graph. The \texttt{LABEL} symbol specifies the label of the resultant node. A label also serves as the name of the node map, which is referred from edge maps.
Therefore, the identifier of each label must be unique in a G2GML description.
Each node pattern contains zero or more properties as \texttt{prop\_list}.
A \texttt{property} is a pair of \texttt{PROP\_NAME} and \texttt{PROP\_VAL} that describes the properties of nodes.
Identifiers \texttt{NODE} and \texttt{PROP\_VAL} must be names of variables in the succeeding RDF graph pattern, while \texttt{LABEL} and \texttt{PROP\_NAME} are regarded as text literals.

A PG edge pattern specifies the pattern of the edge, label, properties, as well as the source and destination nodes.
Here, we define the syntax of the PG edge pattern as follows.
Notably, \texttt{SRC}, \texttt{SRC\_LAB}, \texttt{EDGE}, \texttt{EDGE\_LAB}, \texttt{DST}, and \texttt{DST\_LAB} are all identifiers.
\begin{defi}[EBNF notation of PG edge pattern]
\leavevmode
\begin{verbatim}
edge_pat ::= "(" SRC  ":" SRC_LAB ") -" 
                "[" EDGE?  ":" EDGE_LAB  ("{" prop_list "}")? ")" 
                ("->" | "-") "(" DST ":" DST_LAB ")"
\end{verbatim}
\end{defi}

 Identifiers of \texttt{SRC} and \texttt{DST} specify the variables in the succeeding RDF graph pattern that should be mapped to endpoint nodes of resultant edges.
A \texttt{SRC\_LAB} and a \texttt{DST\_LAB} serve not only as labels of resultant nodes, but also as implicit constraints of the edge map.
%TODO add ref to example section.
The resultant source and destination nodes must match patterns in the corresponding node maps.
For this reason, \texttt{SRC\_LAB} and \texttt{DST\_LAB} must be defined in other node maps.
\texttt{EDGE}, \texttt{EDGE\_LAB}, and \texttt{prop\_list} can be described in the same manner as their counterparts in node maps.

The resultant edges can be either directed or undirected. If we use '\texttt{->}' as delimiter between an edge and a destination part, the edge will be a directed edge.

In node and edge maps, a variable bound to \texttt{PROP\_VAL} may have multiple candidates against a single node or edge. If multiple candidates exist, they are concatenated into an array in the same manner as GROUP\_CONCAT in SPARQL. If \texttt{EDGE} is omitted in an edge map, such candidates are concatenated within groups of tuples of \texttt{SRC} and \texttt{DST}.

%%%%%%%%%%%
\subsection{Property Graph Model}
We define the PG model independent of specific graph database implementations. For the purpose of interoperability, we incorporate differences in PG models, taking into consideration multiple labels or property values for nodes and edges, as well as mixed graphs with both directed and undirected edges. The PG model we redefine here requires the following characteristics:

\begin{itemize}
    \item A PG contains nodes and edges.
    \item Each node and edge can have zero or more labels.
    \item Each node and edge can have properties (key-value pairs).
    \item Each property can have multiple values.
    \item Each edge can be directed or undirected.
\end{itemize}
Formally, we define the PG model as follows.

\begin{defi}[Property Graph Model]
\leavevmode \vspace{1mm} \\
A \emph{property graph} is a tuple
$PG = \langle N, E_d, E_u, S, V, P, e, l_n, l_e, p_n, p_e\rangle$, where:
\begin{enumerate}
    \item $N$ is a set of nodes.
    \item $E_d$ is a set of directed edges.
    \item $E_u$ is a set of undirected edges.
    \item $E$ is a set of edges, where $E = E_d \cup E_u$.
    \item $S$ is a set of strings.
    \item $V$ is a set of values of arbitrary data types.
    \item $P$ is a set of properties. Each property assumes the form $p = \langle k,v \rangle$, where $k \in S$ and $v \in 2^V$.
    \item $e: E \to \langle N \times N \rangle$ is a function that yields the endpoints of each directed or undirected edge (if the edge is directed, the first node is the source, and the second node is the destination).
    \item $l_n : N \to 2^S$ is a function mapping each node to its multiple labels.
    \item $l_e : E \to 2^S$ is a function mapping each edge to its multiple labels.
    \item $p_n : N \to 2^P$ is a function used to assign nodes to their multiple properties.
    \item $p_e : E \to 2^P$ is a function used to assign edges to their multiple properties.
\end{enumerate}
\end{defi}

\subsection{Serialization of Property Graphs}
According to our definition of the PG model, we propose serialization in flat text and JSON. The flat text format (PG format) performs better in terms of human readability and line-oriented processing, while the JSON format (JSON-PG format) is best used for server-client communication.

@@

\begin{itemize}
    \item Each line describes a node or an edge.
    \item All the elements in each line are separated by spaces or tabs.
    \item In each of the node lines, the first column contains the node ID.
    \item In each of the edge lines, the first three columns contain the source node ID, direction, and destination node ID.
    \item Each line can contain an arbitrary number of labels.
    \item Each line can contain an arbitrary number of properties (key-value pairs).
\end{itemize}

@@

Furthermore, we implemented command-line tools to convert formats between the flat PG and JSON-PG, where the latter follows the JSON syntax in addition to the above definition. We further transform them into formats for well-known graph databases such as Neo4j, Oracle Database, and Amazon Neptune. The practical use cases of our tools demonstrate that the proposed data model and formats have the capability to describe PG data used in existing graph databases (see \url{https://github.com/g2glab/pg}).

@@

%%%%%%%%%%%%
\subsection{Minimal Examples}
Figure~\ref{fig:example-g2g} shows the minimal example of G2G mapping from RDF to PG data, representing the following five types of typical mapping.

\begin{itemize}
    \item Resource to node: In line 2-4, the RDF resources with type \texttt{:Person} are mapped into the PG nodes using their IRIs as node IDs.
    \item Datatype property to node property: In line 2-4, the RDF datatype property \texttt{:name} is mapped onto the PG node property key \texttt{name}. The literal objects 'Alice' and 'Bob' are mapped onto the node property values.
    \item Object property to edge: In line 5-6, the RDF object property \texttt{:supervised\_by} is mapped onto the PG edge \texttt{supervised\_by}.
    \item Resource to edge: In line 7-10, the RDF resource with type \texttt{:Email} is mapped onto the PG edge \texttt{emailed}. 
    \item Datatype property to edge property: In line 7-10, the RDF datatype property \texttt{:year} and \texttt{:attachment} are mapped onto the PG edge property \texttt{year} and \texttt{attachment}. The literal objects \texttt{2017} and \texttt{01.pdf} are mapped onto the edge property values.
\end{itemize}

@@

%%%%%%%%%%%
\subsection{Mapping Details}
The G2G mapping based on G2GML is generally intuitive, as shown above; however, there are several discussions regarding the details of mapping.

\subsubsection{Referencing Node Labels}
The node labels in edge maps must be defined in node maps (described in Section 2.2) to reference the conditions. This means that the conditions for defining `Person` are imported into the definition of `supervised\_by` and `emailed` relationships. Therefore, the `supervised\_by` relationship will be generated only between the nodes satisfying the `Person` condition. In this manner, users can retrieve the PG datasets that are intuitively consistent (it is also possible to avoid specifying node labels). In contrast, edge conditions are not inherited to nodes, such that the nodes that do not have any `supervised\_by` nor `emailed` relationships are also generated.

\subsubsection{Multi-Edges}
When a pair of RDF resources has multiple relationships, those will be converted to multiple PG edges instead of a single PG edge with multiple properties. The G2GML in Figure~\ref{fig:example-g2g} generates two PG edges for the RDF data in Figure~\ref{fig:example-rdf2} to maintain the information that Alice emailed Bob twice.

\begin{figure}[!t]
\begin{scriptsize}
\begin{verbatim}
@prefix : <http://example.org/> .
:person1 a :Person ;
         :name 'Alice' .
:person2 a :Person ;
         :name 'Bob' .
[] a :Email ;
   :sender   :person1 ;
   :receiver :person2 ;
   :year     2017 .
[] a :Email ;
   :sender   :person1 ;
   :receiver :person2 ;
   :year     2018 .
\end{verbatim}
\end{scriptsize}
\caption{Example of input RDF data (multi-edges)}
\label{fig:example-rdf2}
\end{figure}

\begin{figure}[!t]
\begin{scriptsize}
\begin{verbatim}
"http://example.org/person1" :Person name:Alice
"http://example.org/person2" :Person name:Bob
"http://example.org/person1" -> "http://example.org/person2" :emailed year:2017
"http://example.org/person1" -> "http://example.org/person2" :emailed year:2018
\end{verbatim}
\end{scriptsize}
\caption{Example of output PG data (multi-edges)}
\label{fig:example-pg2}
\end{figure}

\subsubsection{List of Property Values}
Each PG property can assume multiple values. In an RDF that includes the same data property multiple times, the values are assumed to be the members of a list. The G2GML in Figure~\ref{fig:example-g2g} generates one PG edge with two properties for the RDF data in Figure~\ref{fig:example-rdf3}, keeping the information that Alice emailed Bob once, and the email contained two attachments.
